---
title: "Airbnb_project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
```

# Airbnb listings analysis project

*by Océane Salmeron, Nathan Lancman, Charles Mabbyalas, Baptise Lafay (ING5, BDA 2)*

Airbnb data contains a lot of value.

A real estate investor is looking forward to using this data in order to take a decision regarding the best option he has in real estate to generate profit on Airbnb from his investment.

He downloaded the latest available data at a specific time about the following cities: 

*   Malaga
*   Mallorca
*   Sevilla

He's looking forward to answering the following questions:

```{r}
library(dplyr)
library(stringr)
library(ggplot2)
library(data.table)
library(tidyr)
library(forcats)

setwd("/Users/oceanesalmeron/Documents/Data Analytics/airbnb_listings_analysis")

# a generic function to prepare data for a specific city, data_date




## Once data for multiple cities are prepared
## We can read these data and concatenate them together into one dataframe

# Reading cleansed data
cities <- c("malaga", "mallorca", "sevilla")
data_dates <- c("2020-06-30", "2020-09-19", "2020-06-29")

# We are only interested in data between min_date and max_date
min_date <- '2020-05-01'
max_date <- '2020-11-01'

files_paths <- c()

# Read data in cities between min_date and max_date
for(city in cities){
    file_dir <- file.path(".", "App/Data/data_cleansed", city)
    file_subdirs <- list.dirs(file_dir)
    file_subdirs <- file_subdirs[-1]
    
    for(file_subdir in file_subdirs){
        if(file_subdir < file.path(file_dir, min_date) | file_subdir > file.path(file_dir, max_date)  )
            file_subdirs = file_subdirs[file_subdirs != file_subdir]
    }
    files_paths <- c(files_paths, file_subdirs)
}
#calendars <-
#    do.call(rbind,
#            lapply(file.path(files_paths, "calendar.csv"), read.csv, row.names=1))

files_paths <- file.path(files_paths, "listings.csv")
listings <- 
    do.call(rbind,
            lapply(files_paths, read.csv, row.names=1))



## Preprocess
listings$bedrooms <- ifelse(listings$bedrooms >= 5, "5+", listings$bedrooms)

files_paths
```
# Analysis 1 : Comparing Cities
### 1)Find the "average availability over 30 days" of listings per each city.

```{r}
avg_availability <- listings %>% 
                    group_by(city) %>%
                    summarise(avg = mean(availability_30))

p <-ggplot(avg_availability, aes(x=city, y=avg, fill=city))
p + geom_bar(stat = "identity")+
    ggtitle("Average availability over 30 days of listings per each city")+
    ylab("Average availability over 30 days")+
    xlab("City")+
    guides(fill = FALSE) +
    geom_text(aes(label=avg), vjust=1.6, color="white", size=3.5) +
    scale_fill_brewer(palette = "Set2") +
    theme(plot.title = element_text(hjust = 0.5))

```

### 2)Find the "average revenue of over 30 days" of listings per each city.

```{r}
avg_revenue<- listings %>% 
              group_by(city) %>%
              summarise(avg = mean(revenue_30))

p <-ggplot(avg_revenue, aes(x=city, y=avg, fill=city))
p + geom_bar(stat = "identity")+
    ggtitle("Average revenue over 30 days of listings per each city")+
    ylab("Average revenue over 30 days")+
    xlab("City")+
    guides(fill = FALSE) +
    geom_text(aes(label=avg), vjust=1.6, color="white", size=3.5) +
    scale_fill_brewer(palette = "Set2") +
    theme(plot.title = element_text(hjust = 0.5))
```

### 3)Compare the distribution of estimated availability for the next 30 days of listings per each city.
```{r}
p <- ggplot(listings, aes(city, availability_30, colour=city))
p + geom_boxplot(outlier.shape = NA) +
    scale_y_continuous(limits = quantile(listings$availability_30, c(0.1, 0.9), na.rm = T)) +
    scale_colour_brewer(palette="Set2") +
    ggtitle("Distribution of estimated availability for the next 30 days of listings per each city")+
    ylab("Average avalaibility over 30 days")+
    xlab("City")+
    guides(colour = FALSE) +
    theme(plot.title = element_text(hjust = 0.5))


    
```

### 4)Compare the distribution of estimated revenue for the next 30 days of listings per each city.
```{r}
p <- ggplot(listings, aes(city, revenue_30, colour=city))
p + geom_boxplot(outlier.shape = NA) +
    scale_y_continuous(limits = quantile(listings$revenue_30, c(0.1, 0.9), na.rm = T))+
    stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="red", fill="red") +
    scale_colour_brewer(palette="Set2") +
    ggtitle("Distribution of estimated revenue for the next 30 days of listings per each city")+
    ylab("Estimated revenue for the next 30 days")+
    xlab("City")+
    guides(colour = FALSE) +
    theme(plot.title = element_text(hjust = 0.5))
```

### 5)Compare the distribution of estimated revenue for the next 30 days of listings per each city & for each house size (# of bedrooms).
```{r}
p <- ggplot(listings, aes(bedrooms, revenue_30, colour=city))
p + geom_boxplot(outlier.shape = NA) +
    scale_y_continuous(limits = quantile(listings$revenue_30, c(0.1, 0.9), na.rm = T)) + 
    facet_wrap(~ city,scale="free") +
    scale_colour_brewer(palette="Set2") +
    ggtitle("Distribution of estimated revenue for the next 30 days of listings per each city & for each house size")+
    ylab("Estimated revenue for the next 30 days")+
    guides(colour = FALSE) +
    theme(plot.title = element_text(hjust = 0.5))
  
```

### 6)Compare the distribution of estimated revenue for the next 30 days of listings per each city & for each room type (room_type).

```{r}
p <- ggplot(listings, aes(room_type, revenue_30, colour=city))
p + geom_boxplot(outlier.shape = NA) +
    scale_y_continuous(limits = quantile(listings$revenue_30, c(0.1, 0.9), na.rm = T)) + 
    facet_wrap(~ city,ncol=3) +
    coord_flip()+
    scale_colour_brewer(palette="Set2") +
    ggtitle("Distribution of estimated revenue for the next 30 days of listings per each city & for each room type")+
    ylab("Estimated revenue for the next 30 days")+
    xlab("Room type")+
    guides(colour = FALSE) +
    theme(plot.title = element_text(hjust = 0.5))
  
```


## Analyse 2 : Deep Dive in one City Analysis

### 1)What is the proportion of each room type?

```{r}

p <- ggplot(listings, aes(city,fill=room_type))
p + geom_bar(position ="fill") +
    scale_fill_brewer(palette="Set2") +
    ggtitle("Proportion of each room type by city")+
    ylab("Count")+
    xlab("City")+
    labs(fill="Room type")+
    theme(plot.title = element_text(hjust = 0.5))

```

### 2)hat is the proportion of each house size (# of bedroom)?

```{r}
listings%>%
  filter(!is.na(bedrooms))%>%
  ggplot(aes(city, fill=bedrooms))+
  geom_bar(position = "fill")+
  scale_fill_brewer(palette="Set2") +
  ggtitle("Proportion of each house size by city")+
  ylab("Count")+
  xlab("City")+
  labs(fill="Number of bedroom")+
  theme(plot.title = element_text(hjust = 0.5))

```

### 3)What is the proportion of each neighborhood?

```{r}
listings%>%
  sample_n(size=20)%>%
  ggplot(aes(neighbourhood_cleansed)) + 
  geom_bar(fill="lightgreen") +
  coord_flip()+ 
  facet_wrap(~city, ncol=3)+
  ggtitle("Proportion of each neighborhood by city")+
  ylab("Count")+
  xlab("Neighborhood")+
  theme(plot.title = element_text(hjust = 0.5))
```

### 4)What is the average availability over the next 30 days for each room type / house size / neighborhood?

#### 4.1)Room type
```{r}
avg_availability_room <- listings %>% 
                    group_by(city,room_type) %>%
                    summarise(avg = mean(availability_30))

p <- ggplot(avg_availability_room, aes(fill=room_type, y=avg, x=room_type)) 
p + geom_bar(position="dodge", stat="identity")+
  facet_wrap(~city)+
  scale_fill_brewer(palette="Set2") +
  ggtitle("Average availability over the next 30 days for each room type by city")+
  ylab("Availability over the next 30 days")+
  xlab("")+
  labs(fill="Room type")+
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(),axis.ticks = element_blank())
```

#### 4.2)House size

```{r}
avg_availability_bedrooms <- listings %>% 
                              filter(!is.na(bedrooms))%>%
                              group_by(city,bedrooms) %>%
                              summarise(avg = mean(availability_30))

p <- ggplot(avg_availability_bedrooms, aes(fill=bedrooms, y=avg, x=bedrooms)) 
p + geom_bar(position="dodge", stat="identity")+
  facet_wrap(~city)+
  scale_fill_brewer(palette="Set2") +
  ggtitle("Average availability over the next 30 days for each house size by city")+
  ylab("Availability over the next 30 days")+
  xlab("")+
  labs(fill="Number of bedroom")+
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(),axis.ticks = element_blank())

```

#### 4.3)Neighbourhood

```{r}
avg_availability_neigh <- listings %>% 
                    sample_n(size=20)%>%
                    group_by(city,neighbourhood_cleansed) %>%
                    summarise(avg = mean(availability_30))

p <- ggplot(avg_availability_neigh, aes(fill="red", y=avg, x=neighbourhood_cleansed)) 
p + geom_bar(position="dodge", stat="identity")+
  facet_wrap(~city)+
  coord_flip()+ 
  ggtitle("Average availability over the next 30 days for each neighbourhood by city")+
  ylab("Availability over the next 30 days")+
  xlab("Neighbourhood")+
  guides(fill=FALSE)+
  theme(plot.title = element_text(hjust = 0.5))
```

### 5)What is the average revenue over the next 30 days for each room type / house size / neighborhood?

#### 5.1)Room type
```{r}
avg_revenue_room <- listings %>% 
                    group_by(city,room_type) %>%
                    summarise(avg = mean(revenue_30))

p <- ggplot(avg_revenue_room, aes(fill=room_type, y=avg, x=room_type)) 
p + geom_bar(position="dodge", stat="identity")+
  facet_wrap(~city)+
  scale_fill_brewer(palette="Set2") +
  ggtitle("Average revenue over the next 30 days for each room type by city")+
  ylab("Revenue over the next 30 days")+
  xlab("")+
  labs(fill="Room type")+
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(),axis.ticks = element_blank())
```


#### 5.2)House size
```{r}
avg_revenue_bedrooms <- listings %>% 
                    filter(!is.na(bedrooms))%>%
                    group_by(city,bedrooms) %>%
                    summarise(avg = mean(revenue_30))

p <- ggplot(avg_revenue_bedrooms, aes(fill=bedrooms, y=avg, x=bedrooms)) 
p + geom_bar(position="dodge", stat="identity")+
  facet_wrap(~city)+
  scale_fill_brewer(palette="Set2") +
  ggtitle("Average revenue over the next 30 days for each house size by city")+
  ylab("Revenue over the next 30 days")+
  xlab("")+
  labs(fill="Number of bedroom")+
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(),axis.ticks = element_blank())
```


#### 5.3)Neighbourhood
```{r}
avg_revenue_neigh <- listings %>% 
                    sample_n(size=20)%>%
                    group_by(city,neighbourhood_cleansed) %>%
                    summarise(avg = mean(revenue_30))

p <- ggplot(avg_revenue_neigh, aes(fill="red", y=avg, x=neighbourhood_cleansed)) 
p + geom_bar(position="dodge", stat="identity")+
  facet_wrap(~city)+
  coord_flip()+ 
  ggtitle("Average revenue over the next 30 days for each neighbourhood by city")+
  ylab("Revenue over the next 30 days")+
  xlab("Neighbourhood")+
  guides(fill=FALSE)+
  theme(plot.title = element_text(hjust = 0.5))
```

### 6)What is the distribution of availability over the next 30 days for each room type / house size / neighborhood?

#### 6.1)Room type
```{r}
p <- ggplot(listings, aes(room_type, availability_30))
p + geom_boxplot(aes(colour = city), outlier.shape = NA) +
    scale_y_continuous(limits = quantile(listings$availability_30, c(0.1, 0.9), na.rm = T)) + 
    facet_wrap(~ city,ncol=3)+
    coord_flip()+ 
    scale_colour_brewer(palette="Set2") +
    ggtitle("Distribution of availability over the next 30 days for each room type by city")+
    ylab("Estimated availability for the next 30 days")+
    xlab("Room type")+
    guides(colour = FALSE) +
    theme(plot.title = element_text(hjust = 0.5))
```

#### 6.2)House size
```{r}
listings%>% filter(!is.na(bedrooms))%>%
          ggplot(aes(bedrooms, availability_30)) +
          geom_boxplot(aes(colour = city), outlier.shape = NA) +
          scale_y_continuous(limits = quantile(listings$availability_30, c(0.1, 0.9), na.rm = T)) + 
          facet_wrap(~ city,scale="free")+
          scale_colour_brewer(palette="Set2") +
          ggtitle("Distribution of availability over the next 30 days for house size type by city")+
          ylab("Estimated availability for the next 30 days")+
          xlab("House size")+
          guides(colour = FALSE) +
          theme(plot.title = element_text(hjust = 0.5))
```


#### 6.3)Neighbourhood
```{r}
listings%>%sample_n(size=20)%>%
             ggplot(aes(neighbourhood_cleansed, availability_30))+
             geom_boxplot(aes(colour = city), outlier.shape = NA) +
             scale_y_continuous(limits = quantile(listings$availability_30, c(0.1, 0.9), na.rm = T)) + 
             facet_wrap(~ city,ncol=3)+
             coord_flip()+ 
             scale_colour_brewer(palette="Set2") +
             ggtitle("Distribution of availability over the next 30 days for each neighbourhood by city")+
             ylab("Estimated availability for the next 30 days")+
             xlab("Neighbourhood")+
             guides(colour = FALSE) +
             theme(plot.title = element_text(hjust = 0.5))
  

```


### 7)What is the distribution of revenue over the next 30 days for each room type / house size / neighborhood?

### 7.1)Room type

```{r}
p <- ggplot(listings, aes(room_type, revenue_30))
p + geom_boxplot(aes(colour = city), outlier.shape = NA) +
    scale_y_continuous(limits = quantile(listings$revenue_30, c(0.1, 0.9), na.rm = T)) + 
    facet_wrap(~ city,ncol=3)+
    coord_flip()+ 
    scale_colour_brewer(palette="Set2") +
    ggtitle("Distribution of revenue over the next 30 days for each room type by city")+
    ylab("Estimated revenue for the next 30 days")+
    xlab("Room type")+
    guides(colour = FALSE) +
    theme(plot.title = element_text(hjust = 0.5))
```

#### 7.2)House size
```{r}
listings%>% filter(!is.na(bedrooms))%>%
          ggplot(aes(bedrooms, revenue_30)) +
          geom_boxplot(aes(colour = city), outlier.shape = NA) +
          scale_y_continuous(limits = quantile(listings$revenue_30, c(0.1, 0.9), na.rm = T)) + 
          facet_wrap(~ city,scale="free")+
          scale_colour_brewer(palette="Set2") +
          ggtitle("Distribution of revenue over the next 30 days for house size type by city")+
          ylab("Estimated revenue for the next 30 days")+
          xlab("House size")+
          guides(colour = FALSE) +
          theme(plot.title = element_text(hjust = 0.5))
```


#### 7.3)Neighbourhood
```{r}
listings%>%sample_n(size=20)%>%
            ggplot(aes(neighbourhood_cleansed, revenue_30))+
            geom_boxplot(aes(colour = city), outlier.shape = NA) +
            scale_y_continuous(limits = quantile(listings$revenue_30, c(0.1, 0.9), na.rm = T)) + 
            facet_wrap(~ city,ncol=3)+
            coord_flip()+
            scale_colour_brewer(palette="Set2") +
            ggtitle("Distribution of revenue over the next 30 days for each neighbourhood by city")+
            ylab("Estimated revenue for the next 30 days")+
            xlab("Neighbourhood")+
            guides(colour = FALSE) +
            theme(plot.title = element_text(hjust = 0.5))
```














